### 2024-06-02 player-evaluation.qmd
20240602 - integration of evaluation information

The discussion that follows is based on my and others' *impressions* of which characteristics may be valuable to consider when evaluating players.
However, the extent to which any factor is actually relevant for predicting future performance is an empirical question and should be evaluated empirically.

# Athletic Profile {sec-evalAthleticProfile}

@@

# Historical Performance {sec-evalHistoricalPerformance}

@@

# Health {sec-evalHealth}

@@

# Age and Career Stage {sec-evalAgeCareerStage}

@@

# Situational Factors {sec-evalSituation}

@@

# Matchups {sec-evalMatchups}

@@

# Cognitive and Motivational Factors {sec-evalCogMotivational}

@@

>>>>

The discussion that follows is based on my and others' *impressions* of some of the characteristics that may be valuable to consider when evaluating players.
However, the extent to which any factor is actually relevant for predicting future performance is an empirical question and should be evaluated empirically.

## Athletic Profile {#sec-evalAthleticProfile}

@@

## Historical Performance {#sec-evalHistoricalPerformance}

@@

## Health {#sec-evalHealth}

@@

## Age and Career Stage {#sec-evalAgeCareerStage}

@@

## Situational Factors {#sec-evalSituation}

@@

## Matchups {#sec-evalMatchups}

@@

## Cognitive and Motivational Factors {#sec-evalCogMotivational}

@@

## Putting it Altogether {#sec-evalIntegration}

After performing an evaluation of the relevant domain(s) for a given player, then one must integrate the evaluation information across domains to make a judgment about a player's overall value.
When thinking about a player's value, it can be worth thinking of a player's upside (or ceiling) and a player's downside (or flow).
Player that are more consistent may show higher downside but a lower upside.
Younger, less experienced players may show a higher upside but a lower downside.

The extent to which you prioritize a higher upside versus a higher downside may depend on many factors.
For instance, when drafting players, you may prioritize drafting players with the highest downside (i.e., the safest players), whereas you may draft sleepers (i.e., players with higher upside) for your bench.
When choosing which players to start in a given week, if you are predicted to beat a team handily, it may make sense to start the players with the highest downside.
By contrast, if you are predicted to lose to a team by a good margin, it may make sense to start the players with the highest upside.
